### 2015-06-12 text/ms.Rmd
Add some draft result notes

In fisheries science, a common problem is estimating the status of an exploited
fish population. For the majority of fish stocks, we have limited biological
information and stock status isn't known [@fao2014]. In recent years, a number
of methods have been proposed to derive population status based on the limited
information that is frequently available: fisheries catch time series and basic
knowledge about the productivity of the species. Recently, @rosenberg2014
investigated the ability for these models to estimate population status through
a large-scale simulation model. A key finding was that these models frequently
disagree on population status and any one model has relatively poor predictive
ability (e.g. Fig. \ref{fig:motivate}).

@@

We first developed and tested ensemble models with a fully factorial simulated
dataset [@rosenberg2014]. The simulation models included three life histories: small
pelagic, demersal, large pelagic; three levels of how depleted biomass was at
the start of the dataset: 100%, 70%, and 40% of carrying capacity; and four
harvest dynamics: a constant harvest rate, a harvest rate that is coupled with
biomass to mimic an open-access single-species fishery, a scenario where
harvest rate increases to a fixed level, a 'roller-coaster' scenario where the
harvest rate increases, maintains a steady level and decreases to a lower
steady level. Process noise was introduced to the models through two levels of
multiplicative recruitment variability: $N(0, 0.2)$ and $N(0, 0.6)$, two levels
of first-order autoregressive correlation on recruitment variability: 0 and
0.6. The simulation included a scenario without observation error and with
multiplicative observation error around catch at $N(0, 0.2)$.

@@

- *SSCOM* (state-space catch-only-model) ...

@@

* Individual model and ensemble model scatterplots:
Fig. \ref{fig:hexagon}

* Accuracy, rank correlation, and bias:
Fig. \ref{fig:performance}

* Regression coefficients from linear model ensemble:
Fig. \ref{fig:lm-coefs}

* Partial dependence plots of GBMs:
Fig. \ref{fig:partial-sim}

* Partial residuals version of GBMs:
Fig. \ref{fig:partial-residuals-sim-gbm}

* 2-dimension partial dependence plots:
Fig. \ref{fig:partial-2d-sim}

* ROC curves from simulation ensembles:
Fig. \ref{fig:roc-sim}

* Scatterplots of RAM fits:
Fig. \ref{fig:hexagon-ram}

* Scatterplots of basic ensembles without covariates:
Fig. \ref{fig:hexagon-sim-basic}

* Scatterplots estimating slope:
Fig. \ref{fig:scatter-sim-slope}

* Accuracy, rank correlation, and bias of slope:
Fig. \ref{fig:performance-sim-slope}

>>>>

In fisheries science, a common problem is estimating the status of an exploited
fish population. For the majority of fish stocks, we have limited biological
information and stock status isn't known [@fao2014]. In recent years, a number
of methods have been proposed to derive population status based on the limited
information that is frequently available: fisheries catch time series and basic
knowledge about the productivity of the species. Recently, @rosenberg2014
investigated the ability for these models to estimate population status through
a large-scale simulation model. A key finding was that these models frequently
disagree on population status and any one model has relatively poor predictive
ability on average (e.g. Fig. \ref{fig:motivate}).

@@

We first developed and tested ensemble models with a fully factorial simulated
dataset [@rosenberg2014]. In summary, the simulation models included three life
histories: small pelagic, demersal, large pelagic; three levels of how depleted
biomass was at the start of the dataset compared to carrying capacity: 100%,
70%, and 40% of carrying capacity; and four harvest dynamics: a constant
harvest rate, a harvest rate that is coupled with biomass to mimic an
open-access single-species fishery, a scenario where harvest rate increases to
a fixed level, a 'roller-coaster' scenario where the harvest rate increases,
maintains a steady level and decreases to a lower steady level. Process noise
was introduced to the models through two levels of multiplicative recruitment
variability: $N(0, 0.2)$ and $N(0, 0.6)$, and that process noise was either
uncorrelated or had a first-order autoregressive correlation of 0.6. The
simulation included a scenario without observation error and with
multiplicative observation error around catch at $N(0, 0.2)$.

@@

- *SSCOM* (state-space catch-only-model) TODO: Jim -- can you give a ~2 sentence summary?

@@

CMSY has bimodality (Fig. \ref{fig:hexagon}a)

Applied to the simulated data (Fig. \ref{fig:hexagon}e--h, Fig. \ref{fig:performance}a)...

machine learning models and linear models outperformed all individual models in terms of accuracy (MAPE) and Spearman's rank-order correlation across stocks

ensembles generally had lower bias (MPE) than any individual model (Fig. \ref{fig:performance})

ensemble models generally had better ability to distinguish if simulated stocks were above or below $B/B_\mathrm{MSY}$ = 1$ (Fig. \ref{fig:roc-sim})

ensembles also did better at estimating the slope of \bbmsy\\ in the last five years (Figs. \ref{fig:scatter-sim-slope}, \ref{fig:performance-sim-slope}).

for slope: again, machine learning ensembles best by rank-order correlation and accuracy and had lowest bias (numbers) (Fig. \ref{fig:performance-sim-slope})

Although the additional spectral density covariates improved ensemble fit, performance of the ensembles was only marginally degraded by removing these covariates (Fig. ~\ref{fig:hexagon} vs. Fig. \ref{fig:hexagon-sim-basic})

For example, the linear model ensemble suggests that if both mPRM and SSCOM predict high \bbmsy\\, the true \bbmsy\\ is also high (Fig. \ref{fig:lm-coefs})

The GBM model, also indicates a similar relationship, but suggests a non-linear relationship between these individual models in predicting \bbmsy\\ (Fig. \ref{fig:partial-2d-sim}XX)

Partial dependence plots illustrate the non-linear relationship between the
various individual model predictions and the ensemble predictions (Fig.
\ref{fig:partial-sim}). These integrate across other predictor values...

Ensemble models can exploit the best performance characteristics of individual models. 
For example, if SSCOM predicts a low \bbmsy\\ (less than about 0.X), true \bbmsy\\ is probably low (Fig. \ref{fig:hexagon}c).
But, SSCOM provides less predictive ability on its own when it predicts higher \bbmsy\\. The machine learning ensemble models can exploit this (Fig. \ref{fig:partial-sim}XX). On its own, CMSY contributes the most to predicting \bbmsy\\ when it makes predictions of \bbmsy\\ over one.

Ensemble models trained on the simulated dataset, outperformed any individual model across all performance dimensions when applied to a completely new dataset: the RAM Legacy Stock Assessment database. They even outperform the mPRM method which is trained on the stock assessment database itself (Fig. \ref{fig:hexagon-ram})

Ensembles take advantage of not just the best aspects of individual models, but also the interactions between multiple individual models or the interaction between individual models and other covariates. For example, CMSY becomes more informative about mean \bbmsy\\ when combined with the spectral density measures (Fig. \ref{fig:partial-2d-sim}XX, XX).

Across simulated data mean \bbmsy\\, stock assessment database \bbmsy\\ and
simulated data slope, the ensembles consistently had the best, or among the best levels of accuracy, rank order correlation, and level of bias (Figs \ref{fig:performance}, \ref{fig:performance-sim-slope}).

Some individual models performed nearly as well as the enembles in one or two performance dimensions: e.g. SSCOM does nearly as well as the ensembles at predicting the slope of \bbmsy\\ (Fig. \ref{fig:performance-sim-slope}). But no individual method is consistently nearly as good as the ensembles. SSCOM does more poorly than the ensembles and other individual models in terms of accuracy, rank-order correlation, and bias at estimating the mean \bbmsy\\ for both the simulated and stock-assessment datasets (Fig. \ref{fig:performance).
